Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
Merge branch 'master' of https://github.com/bcpierce00/interfering
* 'master' of https://github.com/bcpierce00/interfering:
  Eager testing file
  work in 4
  Add copy of lazy properties file for eager DI
  Add eager default layout
  Corrections to eager files
  wordsmithing in 5
  Copy lazy files as baseline for revised eager DI
  Move outdated Eager modules

# Conflicts:
#	submission5/core.tex

The term {\em stack safety} is associated with a
variety of compiler, run-time, and hardware mechanisms for protecting stack
memory. But, unlike ``the heap,'' the stack does not correspond to a single
high-level language concept. Rather, it corresponds to the fundamental
abstraction of functions, in all of the forms that they can take.
This protean nature of the functional abstraction makes stack safety
difficult to specify.

We propose a formal characterization of stack safety based on
concepts from language-based security. Stack safety is decomposed
into an integrity property and a confidentiality property for each
of the caller and the callee, as well as a control-flow property.

Our motivating enforcement mechanisms,
the ``lazy'' stack safety micro-policies proposed by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, permit functions
to write into one another's frames, but taint the result so that the frame's
owner cannot access it. No existing characterization of stack safety
captures this style of safety. We capture it by defining the properties in
terms of the observable behavior of the system.

The stack interacts with a large number of language features that are often excluded
from discussion of stack safety. Our properties support a system with
both caller- and callee-saved registers, arguments passed on the stack,
\ifexceptions tail-call elimination, and exceptions.
\else and tail-call elimination. \fi They are modular by
design, in order to be further extensible to other features.

We validate our properties by using them to distinguish correct implementations
of Roessler and DeHon's micro-policies from incorrect ones
via property-based random testing. Our testing successfully detects violations
in several broken variants, including Roessler and DeHon's original lazy policy.
A fixed version of the policy does pass our tests.

>>>>

The term {\em stack safety} is used for a variety of compiler,
run-time, and hardware mechanisms for protecting stack memory. Unlike
``the heap,'' the hardware stack does not correspond to a single
high-level language concept: different compilers use it in different
ways to support procedural and functional abstraction mechanisms from
a wide range of languages.  This protean nature makes it difficult to
nail down what it means to correctly enforce stack safety.

We propose a formal characterization of stack safety using concepts
from language-based security. Rather than packaging all aspects of
stack safety into a monolithic property, we decompose it into an
integrity property and a confidentiality property for each of the
caller and the callee, plus a control-flow property---five properties
in all.

This presentation is motivated by a particular class of enforcement
mechanisms, the ``lazy'' stack safety micro-policies studied by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, which permit
functions to write into one another's frames, but which taint the
changed locations so that the frame's owner cannot access it. No
existing characterization of stack safety captures this style of
safety. We capture it here by stating our properties in terms of the
observable behavior of the system.

The hardware stack interacts with a large number of language features,
some of which are typically excluded from discussions of stack
safety. Our properties go further than previous proposals in
supporting caller- and callee-saved registers, arguments passed on the
stack, \ifexceptions tail-call elimination, and exceptions.  \else and
tail-call elimination. \fi They are modular by design, in order to be
further extensible to other features.

We validate our properties by using them to distinguish between
correct and incorrect implementations of Roessler and DeHon's
micro-policies using property-based random testing. Our test harness
successfully identifies several broken variants, including Roessler
and DeHon's original lazy policy.  A fixed version of their policy
does pass our tests.
